---
title: "1-ipipneo120"
author: "Johannes Titz"
date: "12/10/2021"
output:
  html_document:
    toc: true
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE, message = FALSE,
                      warning = FALSE)
```

## Intro

This is analysis 1 of the paper. First, install the `star` package:

```{r}
if (!require("star")) {
  install.packages("star_0.1.0.tar.gz", repos = NULL, type = "source")
}
```

Load some packages. I will use librarian for installing/loading packages. 

```{r}
if (!require('librarian')) install.packages('librarian')
librarian::shelf(tidyr, dplyr, ggplot2, cowplot, corrgram, ggraph, igraph,
                 GGally, network, ggnetwork, tidygraph, furrr, star,
                 progressr, simpleCache, lavaanPlot, lavaan, DiagrammeRsvg,
                 rsvg)
```

Helper function

```{r}
get_stats <- function(var, df) {
means <- df %>%
  group_by(.data[[var]]) %>%
  summarize(mean = mean(sum_score, na.rm = TRUE),
            se = sd(sum_score, na.rm = TRUE) / length(na.omit(sum_score)))
  names(means)[1] <- "group"
  means
}
```

Set cache dir for simpleCache:

```{r}
setCacheDir("cache")
```

load labels for ipip

```{r}
labels <- create_ipip_neo_items()
```

The original data is available at https://osf.io/tbmh5/ It is too large to make it available through `data`, but `star` has the data set under inst/extdata:

```{r data}
fpath <- system.file("extdata", "ipipneo.rda", package = "star")
if (!exists("ipipneo")) load(fpath)
```

Set simulation parameters. I am on GNU/Linux, please check that the detection of cores works for you. If it does not work, just set it to a fixed value. `star` uses `parallel` for parallel computation. Sorry, Windows users, parallelization only works under GNU/Linux. You should still be able to run the code, but you cannot take advantage of multiple cores. I tried to implement parallelization via futures (`furrr`), but this turned out to be slower than using a single core.

```{r params}
mc_cores <- round(parallel::detectCores() - 1)
alpha <- 0.05 / (24 * 23 * 5)
param <- sim_parameters(n_bootstrap_samples = 10, mc_cores = mc_cores)
```

## first analysis, with only 10 simulation runs, only keep p-values of 0

```{r all10}
set.seed(5989)
simpleCache(
  "ipipneo10",
  {test_items_against_dimension(names(ipipneo), ipipneo, param)},
  timer = TRUE
)
left_items <- extract_sig_items(ipipneo10, 0)
```

Check how many items could not be tested properly:

```{r problems}
ipipneo10 %>%
  group_by(i) %>%
  summarize(warn = mean(var0 + n_min_violated > 0)) %>%
  filter(warn > 0)
```

None.

## Stepwise tests with more bootstrap samples

The code is not pretty, maybe accumulate can be used here?

```{r stepwise}
steps <- c(1e2, 1e3, 1e4, 1e5)
result <- ipipneo10
for (step in steps) {
  param$n_bootstrap_samples <- step
  simpleCache(
    paste0("stepwise", step),
    {test_specific_items(ipipneo, result$i, result$j, param)},
    timer = TRUE,
    assignTo = "result",
  )
  result <- result[result$p <= alpha, ]
}
result$group <- substr(result$i, 1, 1)
result <- left_join(result, labels, by = c("i" = "item_char"))
# a better deviance value for network graph
result <- result %>%
  mutate(deviance_norm = sqrt(deviance_data / n) * 100)
# display result
result %>%
  select(i, j, p, deviance_data) %>%
  filter(p <= alpha)
```

## some exemplary severe violations

First change item labels for consistency (capitalize):

```{r}
result <- mutate(result,
                 i = stringr::str_to_upper(i),
                 j = stringr::str_to_upper(j))
names(ipipneo) <- stringr::str_to_upper(names(ipipneo))
```

```{r}
data(ipipneo_items)
items <- ipipneo_items
severe <- arrange(result, -deviance_data) %>%
  group_by(group) %>%
  slice(1:2) %>% ungroup() %>% slice(1, 3, 5, 7, 9, 10)
severe <- severe %>% mutate(item_number_i = as.numeric(substr(i, 2, 4)),
                            item_number_j = as.numeric(substr(j, 2, 4)))
severe <- left_join(severe, items, by = c("item_number_i" = "item_number"))
severe <- left_join(severe, items, by = c("item_number_j" = "item_number"))
severe <- mutate(severe, item_content.x = paste(i, item_content.x),
                 item_content.y = paste(j, item_content.y))
  # slice(1:6, 143, 144, 285, 286)
plots <- Map(function(x, y) plot_it(ipipneo[, c(x, y)], with_n = FALSE), severe$i, severe$j)

plots2 <- Map(function(plot, item_x, item_y) plot + scale_x_continuous(item_x) +
      scale_y_continuous(item_y), plots, severe$item_content.x, severe$item_content.y)
plots2
pdf("plots/ipip_neo_severe.pdf", width = 8.5 * 1.1, height =  10 * 1.1)
plot_grid(plotlist = plots2, ncol = 2, labels = "AUTO", label_x = 0.52)
dev.off()
```

## network graph

This code is somewhat messy.

```{r}
sums <- result %>%
  group_by(i) %>%
  summarize(deviance_sum = sum(deviance_data)) %>%
  arrange(desc(deviance_sum))
```

```{r}
## network graph -----
vertices <- data.frame(name = unique(c(result$i, result$j)))
vertices$factor <- substr(vertices$name, 1, 1)
vertices$factor_num <- match(vertices$factor,
                             names(c("N" = 1, "E" = 2, "O" = 3, "A" = 4, "C" = 5)))
vertices$facet <- (as.numeric(substr(vertices$name, 2, 4)) %% 30 - vertices$factor_num + 5 ) / 5
vertices$facet <- ifelse(vertices$facet == 0, 6, vertices$facet)
vertices_a <- vertices[vertices$factor != "O", ]

graph1 <- as_tbl_graph(
  graph_from_data_frame(result[!grepl("O", result$i), c("i", "j", "p", "deviance_data", "deviance_norm")],  directed = TRUE,
                        vertices = vertices_a)
)
graph1

ggraph1 <- ggraph(graph1, layout = "kk") +
  #geom_edge_link(arrow = arrow(length = unit(2, "mm")), end_cap = circle(2, "mm")) +
  geom_edge_link(aes(start_cap = label_rect(node1.name),
                     end_cap = label_rect(node2.name), alpha = 1,
                     show_guide = FALSE,
                     width = deviance_norm),# + #,
                 arrow = arrow(length = unit(2, 'mm')),
                 check_overlap = T) +
  scale_edge_width(range = c(0, 3), name  = "normalized deviance",
                   breaks = seq(0, 18, 1), guide = F) + 
  scale_edge_alpha(guide = "none") +
  #geom_node_point(shape = 21, size = 10) +
  #geom_node_text(aes(label = name), color = "black") +
  geom_node_label(aes(label = name, fill = as.factor(facet)), alpha = 0.5) +
  scale_fill_brewer(palette = "Set1", name = "Facet", position = "bottom") +
                  #fill = "lightgrey") +
  theme_graph() +
  theme(plot.margin = unit(c(0.2, 0.1, 0.05, 0.1), "cm"), legend.position = "bottom")
ggraph1
```

```{r}
graph2 <- as_tbl_graph(
  graph_from_data_frame(result[grepl("O", result$i), c("i", "j", "p", "deviance_data", "deviance_norm")],  directed = TRUE,
                        vertices = vertices[vertices$factor == "O", ]),
)
graph2

ggraph2 <- ggraph(graph2, layout = "kk") +
  #geom_edge_link(arrow = arrow(length = unit(2, "mm")), end_cap = circle(2, "mm")) +
  geom_edge_link(aes(start_cap = label_rect(node1.name),
                     end_cap = label_rect(node2.name), alpha = 1,
                     width = deviance_norm), 
                 arrow = arrow(length = unit(2, 'mm'))) +
  scale_edge_width(range = c(0, 3), name  = "normalized deviance",
                       breaks = seq(0, 18, 1), guide = F) + 
  #geom_node_point(shape = 21, size = 10) +
  #geom_node_text(aes(label = name), color = "black") +
  geom_node_label(aes(label = name, fill = as.factor(facet)), alpha = 0.5) +
  scale_fill_brewer(palette = "Set1", name = "Facet") +
  theme_graph() +
  theme(plot.margin = unit(c(0.05, 0.1, 0.2, 0.1), "cm"), legend.position = "none")
ggraph2
```

```{r}
cairo_pdf("plots/ipip_neo_network_all.pdf", width = 10, height = 14)
plot_grid(ggraph1, ggraph2, nrow = 2)
dev.off()
```

## power analysis for items N51, A84, and liberalism/intellect

```{r powern51, eval = T}
i <- rep("N51", 4)
j <- c("N16", "N46", "N76", "N106")

power51 <- create_object(
  ipipneo, i, j, sample_size = 857,
  bootstrap_samples_power = 1e3,
  bootstrap_samples_monoreg = 1e3,
  alpha = 0.05/3
)
simpleCache("powern51", {calculate_power(power51)$power}, reload = T)
powern51
```

```{r powera84, eval = T}
i <- rep("A84", 7)
j <- c("A4", "A34", "A64", "A94", "A99", "A74", "A49")

res <- create_object(
  ipipneo, i, j, sample_size = 857,
  bootstrap_samples_power = 1e3,
  bootstrap_samples_monoreg = 1e3,
  alpha = 0.05/3
)
simpleCache("powera84", {calculate_power(res)$power})
powera84
```

```{r liberalism}
o6 <- labels %>%
  filter(dimension == "O", facet == 6) %>%
  pull(item_char)
o5 <- labels %>%
  filter(dimension == "O", facet == 5) %>%
  pull(item_char)

grid <- expand.grid(o6, o5, stringsAsFactors = F)

res <- create_object(
  ipipneo, grid$Var1, grid$Var2, sample_size = 857,
  bootstrap_samples_power = 1e3,
  bootstrap_samples_monoreg = 1e3,
  alpha = 0.05/3
)
simpleCache("power_liberalism", {calculate_power(res)$power})
power_liberalism
```

## most problematic items

```{r}
result %>%
  group_by(i) %>%
  summarize(sum_dev = sum(deviance_data, na.rm = TRUE)) %>%
  arrange(sum_dev)
```

### c110

One of the most interesting items from the network graph is C110 because it is inconsistent with all other 23 items.

```{r}
c <- ipipneo %>%
  select(starts_with("C") & !starts_with("C110"))
sum_score <- rowSums(c)
d <- data.frame(C110 = ipipneo$C110, sum_score)
p <- ggplot(d, aes(C110, sum_score)) + 
  stat_summary(fun.data = "mean_cl_normal",
               fun.args = list(conf.int = .95)) +
                     theme_classic() +
  scale_x_continuous("C110 I put little time and effort into my work. (R)") +
  scale_y_continuous("Sum score conscientiousness")
p
cairo_pdf("plots/c110.pdf", width = 5, height = 4)
p
dev.off()
```

sem

```{r}
m0 <- "
  # latent variables
  selfefficacy =~ C5 + C35 + C65 + C95
  orderliness =~ C10 + C40 + C70 + C100
  con3 =~ C15 + C45 + C75 + C105
  con4 =~ C20 + C50 + C80 + C110
  con5 =~ C25 + C55 + C85 + C115
  con6 =~ C30 + C60 + C90 + C120
  c =~ selfefficacy + orderliness + con3 + con4 + con5 + con6
"

ipipneo2 <- ipipneo
ipipneo2$C110[ipipneo2$C110 == 1] <- NA

fit0 <- sem(model = m0, data = ipipneo)
fit1 <- sem(model = m0, data = ipipneo2)

sol0 <- standardizedsolution(fit0, type = "std.all")
sol1 <- standardizedsolution(fit1, type = "std.all")
# facet on C
sol0 %>% filter(lhs == "c", rhs == "con4") %>%
  select(lhs, rhs, est.std)
# item on facet
sol0 %>% filter(lhs == "con4", rhs == "C110") %>%
  select(lhs, rhs, est.std)
# item on facet after "fix"
sol1 %>% filter(lhs == "con4", rhs == "C110") %>%
  select(lhs, rhs, est.std)
```

## item E107

```{r}
e5 <- labels$item_char[labels$dimension == "E" & labels$facet == 5]
sum_score <- ipipneo %>% 
  select(e5) %>% 
  rowSums(na.rm = TRUE)
df <- data.frame(E107 = ipipneo$E107, sum_score)
means <- df %>%
  group_by(E107) %>%
  summarize(mean = mean(sum_score, na.rm = TRUE), sd = sd(sum_score, na.rm = TRUE), n())

myplot <- ggplot(means, aes(E107, mean)) + 
  geom_point(shape = 20, size = 5) + 
  theme_classic() + 
  scale_y_continuous("Sum score excitement seeking") + 
  scale_x_continuous("E107 I like to take it easy (R)")
myplot
cairo_pdf("plots/e107.pdf", width = 5, height = 4)
myplot
dev.off()
cor(df, use = "complete.obs")
cor(ipipneo$E107, ipipneo[, e5], use = "complete.obs")
cors <- sort(cor(ipipneo$E107, ipipneo[, grepl("^E", names(ipipneo))], use = "complete.obs"))
mean(cors)
```

### e107 sem

```{r}
m0 <- "
  # latent variables
  ex1 =~ E2 +  E32 + E62 + E92
  ex2 =~ E7 +  E37 + E67 + E97
  ex3 =~ E12 + E42 + E72 + E102
  ex4 =~ E17 + E47 + E77 + E107
  ex5 =~ E22 + E52 + E82 + E112
  ex6 =~ E27 + E57 + E87 + E117
  extraversion =~ ex1 + ex2 + ex3 + ex4 + ex5 + ex6
"

m1 <- "
  # latent variables
  ex1 =~ E2 +  E32 + E62 + E92
  ex2 =~ E7 +  E37 + E67 + E97
  ex3 =~ E12 + E42 + E72 + E102
  ex4 =~ E17 + E47 + E77
  ex5 =~ E22 + E52 + E82 + E112
  ex6 =~ E27 + E57 + E87 + E117
  extraversion =~ ex1 + ex2 + ex3 + ex4 + ex5 + ex6
"

fit0 <- sem(model = m0, data = ipipneo)
fit1 <- sem(model = m1, data = ipipneo)

sol0 <- standardizedsolution(fit0, type = "std.all")
sol1 <- standardizedsolution(fit1, type = "std.all")
```

## Plot liberalism vs. intellect

```{r}
sum_score <- ipipneo %>% select(all_of(o5)) %>% rowSums(na.rm = TRUE)
sum_score_o6 <- ipipneo %>% select(all_of(o6)) %>% rowSums(na.rm = TRUE)

df <- data.frame(o6 = sum_score_o6, o5 = sum_score)

table(df$o6)
df <- df %>%
  # below 3 there are very few values
  filter(o6 >=3)
myplot <- ggplot(df, aes(o6, o5)) +
  stat_summary(fun.data = "mean_cl_normal",
               fun.args = list(conf.int = .95)) +
                     theme_classic() + 
                     scale_x_continuous("Facet liberalism") +
                     scale_y_continuous("Facet intellect")
myplot
cor(df, use = "complete.obs")
cairo_pdf("plots/liberalism_intellect2.pdf", width = 5, height = 4)
myplot
dev.off()
```

## Facets modesty and trust

SEMs:

```{r modestysem}

m0 <- "
  # latent variables
  trust =~ A4 + A34 + A64 + A94
  morality =~ A9 + A39 + A69 + A99
  altruism =~ A14 + A44 + A74 + A104
  cooperation =~ A19 + A49 + A79 + A109
  modesty =~ A24 + A114 + A54 + A84
  sympathy =~ A29 + A59 + A89 + A119
  agreeableness =~ modesty + sympathy + trust + morality + altruism + cooperation
"

m1 <- "
  # latent variables
  trust =~ A4 + A34 + A64 + A94
  morality =~ A9 + A39 + A69 + A99
  altruism =~ A14 + A44 + A74 + A104
  cooperation =~ A19 + A49 + A79 + A109
  modesty =~ A24 + A114
  sympathy =~ A29 + A59 + A89 + A119
  agreeableness =~ modesty + sympathy + trust + morality + altruism + cooperation
"

fit0 <- sem(model = m0, data = ipipneo)
fit1 <- sem(model = m1, data = ipipneo)

plot0 <- lavaanPlot(model = fit0,
                    graph_options = list(overlap = "true", fontsize = "10",
                                         rankdir = "LR"),
                    node_options = list(shape = "box", fontsize = "10"),
                    edge_options = list(color = "grey"),
                    coefs = TRUE, stand = TRUE)
plot1 <- lavaanPlot(model = fit1,
                    graph_options = list(overlap = "true", fontsize = "10",
                                         rankdir = "LR"),
                    node_options = list(shape = "box", fontsize = "10"),
                    edge_options = list(color = "grey"),
                    coefs = TRUE, stand = TRUE)
embed_plot_pdf(plot0, "plots/modesty_sem0.pdf")
embed_plot_pdf(plot1, "plots/modesty_sem1.pdf")

sol0 <- standardizedsolution(fit0, type = "std.all")
sol1 <- standardizedsolution(fit1, type = "std.all")

facet_factor <- sol0 %>%
  filter(lhs == "agreeableness", rhs != "agreeableness") %>%
  select(lhs, rhs, est.std) %>%
  as.data.frame()
facet_factor1 <- sol1 %>%
  filter(lhs == "agreeableness", rhs != "agreeableness") %>%
  select(lhs, rhs, est.std) %>%
  as.data.frame()
tbl <- cbind(facet_factor, est.std2 = facet_factor1$est.std)
tbl %>%
  mutate(diff = est.std2 - est.std)
r2 <- sum(tbl$est.std**2) / 6
r2b <- sum(tbl$est.std2**2) / 6
r2b - r2
colnames(tbl) <- c("lhs", "Facet", "Loading", "Loading w/o A54, A84")
print(xtable::xtable(tbl[, -1], caption = "", label = "tab:modesty"),
                     booktabs = T,
    include.rownames = F, file = "tables/modesty.tex")
```

```{r}
a1_names <- labels %>%
  filter(dimension == "A", facet == 1) %>%
  pull(item_char)

a1_sum_score <- ipipneo %>%
  select(a1_names) %>%
  rowSums(na.rm = TRUE)

a5_names <- labels %>%
  filter(dimension == "A", facet == 5) %>%
  pull(item_char)

a5_sum_score <- ipipneo %>%
  select(a5_names) %>%
  rowSums(na.rm = TRUE)

df <- data.frame(ipipneo[, a5_names], sum_score = a1_sum_score)

res <- lapply(a5_names, get_stats, df)
names(res) <- a5_names
res2 <- plyr::ldply(res, "data.frame")
names(res2)[1] <- "Item"
res2$Item <- factor(res2$Item, levels = a5_names)
mplot <- ggplot(res2, aes(group, mean, ymin = mean - se * qnorm(0.99),
                          ymax = mean + se * qnorm(0.99),
                          group = Item, color = Item, shape = Item)) +
    geom_pointrange() +
    geom_line(linetype = "dashed") +
    scale_x_continuous("Item answer") +
    scale_y_continuous("Sum score trust") +
    scale_color_brewer(type = "div", palette = "Dark2") + 
    theme_classic() + 
    theme(legend.position = c(0.5, 0.25))
cor(df, use = "complete.obs")
mplot

cairo_pdf("plots/modestytrust.pdf", width = 5, height = 4)
mplot
dev.off()
```

## n46

n46, 0.37 5 facot rmodel ladung.

```{r}
n5 <- labels$item_char[labels$dimension == "N" & labels$facet == 5]
sum_score <- ipipneo %>% select(n5) %>% rowSums(na.rm = TRUE)
df <- data.frame(N46 = ipipneo$N46, sum_score)
means <- df %>%
  group_by(N46) %>%
  summarize(mean = mean(sum_score, na.rm = TRUE), sd = sd(sum_score, na.rm = TRUE), n())
ggplot(means, aes(N46, mean)) + 
  geom_point() + 
  theme_classic()
cor(df, use = "complete.obs")
```

```{r}
e_viol <- result %>% filter(i == "N46") %>%
  select(j) %>% unlist()

e <- ipipneo %>%
  select(starts_with("N"))
means <- e %>% group_by(N46) %>%
  summarize_all(mean, na.rm = TRUE)
means_long <- tidyr::pivot_longer(means, cols = -N46)
means_long$viol <- ifelse(means_long$name %in% e_viol, 1, 0)

means_long <- dplyr::left_join(means_long, labels, by = c("name" = "item_char"))

means_long <- means_long %>% filter(viol == 1)
ggplot(means_long, aes(N46, value, group = name, color = as.factor(name))) + geom_point() + geom_line(alpha = 0.25) +
  #facet_grid(~viol) +
  theme_classic()
```

