---
title: "1-ipipneo120"
author: "Johannes Titz"
date: "12/10/2021"
output:
  html_document:
    toc: true
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = F, message = F, warning = F)
```

## Intro

This is analysis 1 of the paper. First, install the `star` package:

```{r eval=TRUE, include=TRUE}
install.packages("star_0.1.0.tar.gz", repos = NULL, type = "source")
```

Load some packages. I will use librarian for installing/loading packages. First line checks if librarian is installed, if not, it is installed.

```{r}
if (!require('librarian')) install.packages('librarian')
librarian::shelf(tidyr, dplyr, ggplot2, cowplot, corrgram, ggraph, igraph,
                 GGally, network, ggnetwork, tidygraph, furrr, star,
                 progressr)
```

The original data is available at https://osf.io/tbmh5/ It is too large to make it available through `data`, but `star` has the data set under inst/extdata:

```{r data}
fpath <- system.file("extdata", "ipipneo.rda", package = "star")
load(fpath)
```

Set simulation parameters. I am on GNU/Linux, please check that the detection of cores works for you. If it does not work, just set it to a fixed value. `star` uses `parallel` for parallel computation. Sorry, Windows users, parallelization only works under GNU/Linux. You should still be able to run the code, but you cannot take advantage of multiple cores. I tried to implement parallelization via futures (`furrr`), but this turned out to be slower than using a single core.

```{r params}
mc_cores <- round(parallel::detectCores() - 1)
alpha <- 0.05 / (24 * 23 * 5)
param <- sim_parameters(n_bootstrap_samples = 10, mc_cores = mc_cores)
```

## first analysis, with only 10 simulation runs, only keep p-values of 0

```{r all10}
ipipneo10 <- xfun::cache_rds(test_items_against_dimension(names(ipipneo), ipipneo, param))
left_items <- extract_sig_items(ipipneo10, 0)
```

Check how many items could not be tested properly:

```{r problems}
ipipneo10 %>% 
  group_by(i) %>% 
  summarize(warn = mean(var0 + n_min_violated > 0)) %>% 
  filter(warn > 0)
```

None.

## Stepwise tests with more bootstrap samples

The code is not pretty, maybe accumulate can be used here?

```{r stepwise}
steps <- c(1e2, 1e3, 1e4, 1e5)
result <- ipipneo10
for (step in steps) {
  param$n_bootstrap_samples <- step
  result <- xfun::cache_rds(test_specific_items(ipipneo, result$i, result$j, param),
                             file = paste0("stepwise", step))
  result <- result[result$p <= alpha, ]
}
result$group <- substr(result$i, 1, 1)
result %>% select(i, j, p, deviance_data) %>% filter(p <= alpha)
```
## items that need to be removed

This code needs to be improved. It would be better to use a network algorithm to remove the problematic items, but it is also fairly easy to see with the naked eye, except for openness.

```{r}
e_rm <- c("e82", "e107", "e112")
e <- filter(result, grepl("e", i) | grepl("e", j))
e <- filter(e, !(i %in% e_rm), !(j %in% e_rm))
e

c_rm <- c("c110", "c75")
c <- filter(result, grepl("c", i) | grepl("c", j))
c <- filter(c, !(i %in% c_rm), !(j %in% c_rm))
c

a_rm <- c("a54", "a84", "a104", "a114", "a24")
a <- filter(result, grepl("a", i) | grepl("a", j))
a <- filter(a, !(i %in% a_rm), !(j %in% a_rm))
a

o <- filter(result, grepl("o", i) | grepl("o", j))
remove <-  names(sort(table(c(o$i, o$j)), decreasing = T))[1:16]
o <- filter(o, !(i %in% remove), !(j %in% remove))
o
d <- create_ipip_neo_items()
d %>% filter(item_char %in% stringr::str_to_upper(remove)) %>%
  arrange(facet, item_number)
```

## some exemplary severe violations

First change item labels for consistency (capitalize):

```{r}
result <- mutate(result,
                 i = stringr::str_to_upper(i),
                 j = stringr::str_to_upper(j))
names(ipipneo) <- stringr::str_to_upper(names(ipipneo))
```

```{r}
data(ipipneo_items)
items <- ipipneo_items
severe <- arrange(result, -deviance_data) %>%
  group_by(group) %>%
  slice(1:2) %>% ungroup() %>% slice(1, 3, 5, 7, 9, 10)
severe <- severe %>% mutate(item_number_i = as.numeric(substr(i, 2, 4)),
                            item_number_j = as.numeric(substr(j, 2, 4)))
severe <- left_join(severe, items, by = c("item_number_i" = "item_number"))
severe <- left_join(severe, items, by = c("item_number_j" = "item_number"))
severe <- mutate(severe, item_content.x = paste(i, item_content.x),
                 item_content.y = paste(j, item_content.y))
  # slice(1:6, 143, 144, 285, 286)
plots <- Map(function(x, y) plot_it(ipipneo[, c(x, y)], with_n = FALSE), severe$i, severe$j)

plots2 <- Map(function(plot, item_x, item_y) plot + scale_x_continuous(item_x) + 
      scale_y_continuous(item_y), plots, severe$item_content.x, severe$item_content.y)
plots2
pdf("plots/ipip_neo_severe.pdf", width = 8.5 * 1.1, height =  10 * 1.1)
plot_grid(plotlist = plots2, ncol = 2, labels = "AUTO", label_x = 0.52)
dev.off()
```

## network graph

This code is somewhat messy.

```{r}
## network graph -----
vertices <- data.frame(name = unique(c(result$i, result$j)))
vertices$factor <- substr(vertices$name, 1, 1)
vertices$factor_num <- match(vertices$factor,
                             names(c("N" = 1, "E" = 2, "O" = 3, "A" = 4, "C" = 5)))
vertices$facet <- (as.numeric(substr(vertices$name, 2, 4)) %% 30 - vertices$factor_num + 5 ) / 5
vertices$facet <- ifelse(vertices$facet == 0, 6, vertices$facet)
vertices_a <- vertices[vertices$factor != "O", ]

graph1 <- as_tbl_graph(
  graph_from_data_frame(result[!grepl("o", result$group), c(1, 2, 4, 5)],  directed = T,
                        vertices = vertices_a)
)
graph1

ggraph1 <- ggraph(graph1, layout = "kk") + 
  #geom_edge_link(arrow = arrow(length = unit(2, "mm")), end_cap = circle(2, "mm")) + 
  geom_edge_link(aes(start_cap = label_rect(node1.name),
                     end_cap = label_rect(node2.name), alpha = 0.5, show_guide = F),
                     #width = critical_fit), 
                 arrow = arrow(length = unit(1, 'mm'))) + 
  scale_edge_alpha(guide = "none") +
  #geom_node_point(shape = 21, size = 10) + 
  #geom_node_text(aes(label = name), color = "black") +
  geom_node_label(aes(label = name, fill = as.factor(facet)), alpha = 0.5) + 
  scale_fill_brewer(palette = "Set3", name = "Facet", position = "bottom") +
                  #fill = "lightgrey") +
  theme_graph() +
  theme(plot.margin = unit(c(0.2, 0.1, 0.05, 0.1), "cm"), legend.position = "bottom")
ggraph1
```

```{r}
graph2 <- as_tbl_graph(
  graph_from_data_frame(result[grepl("o", result$group), c(1, 2, 4:5)],
                        vertices = vertices[vertices$factor == "O", ]),
  directed = T
)
graph2

ggraph2 <- ggraph(graph2, layout = "kk") + 
  #geom_edge_link(arrow = arrow(length = unit(2, "mm")), end_cap = circle(2, "mm")) + 
  geom_edge_link(aes(start_cap = label_rect(node1.name),
                     end_cap = label_rect(node2.name), alpha = .5),
                     #width = critical_fit), 
                 arrow = arrow(length = unit(1, 'mm'))) + 
  #geom_node_point(shape = 21, size = 10) + 
  #geom_node_text(aes(label = name), color = "black") +
  geom_node_label(aes(label = name, fill = as.factor(facet)), alpha = 0.5) + 
  scale_fill_brewer(palette = "Set3", name = "Facet") +
  theme_graph() + 
  theme(plot.margin = unit(c(0.05, 0.1, 0.2, 0.1), "cm"), legend.position = "none")
ggraph2
#facet_nodes(~popularity)

```

```{r}
cairo_pdf("plots/ipip_neo_network_all.pdf", width = 10, height = 14)
plot_grid(ggraph1, ggraph2, nrow = 2)
dev.off()
```

## power analysis for items N51 and A84

```{r powern51}
i <- rep("N51", 4)
j <- c("N16", "N46", "N76", "N106")

power51 <- create_object(
  ipipneo, i, j, sample_size = 857,
  bootstrap_samples_power = 1e3,
  bootstrap_samples_monoreg = 1e3,
  alpha = 0.025
)

power <- xfun::cache_rds(calculate_power(power51)$power)
power
```

```{r powera84}
i <- rep("A84", 7)
j <- c("A4", "A34", "A64", "A94", "A99", "A74", "A49")

res <- create_object(
  ipipneo, i, j, sample_size = 857, 
  bootstrap_samples_power = 1e3,
  bootstrap_samples_monoreg = 1e3,
  alpha = 0.025
)
power <- xfun::cache_rds(calculate_power(res)$power)
power
```