---
title: "NEO-PI-R Combined"
author: "Johannes Titz"
date: "01/12/2023"
output:
  html_document:
    toc: true
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = F, message = F, warning = FALSE)
```

## Intro

This is a new analysis for the revision. If you have not yet installed the `star` package, do it now:

```{r eval=TRUE, include=TRUE}
if (require("star") == F) {
  install.packages("star_0.1.0.tar.gz", repos = NULL, type = "source")
}
librarian::shelf(dplyr, ggplot2)
```

One of the original data is available at https://dataverse.harvard.edu/dataset.xhtml?persistentId=doi:10.7910/DVN/HE6LJR but I also packaged it into `star`:

```{r}
library(star)
data("neo_pi_r_eugene")
neo <- neo_pi_r_eugene
```

The data set contains some participants that produced many missing values:

```{r}
missing <- rowSums(is.na(neo))
stem(missing)
```

Participants with more than 5% missing values were removed (about 2% of the sample).

```{r}
filter <- missing <= nrow(neo) * 0.05
mean(filter)
neo <- neo[filter, ]
```

The other original data set is available in the `qgraph` R package. I restructured the data slightly and repackaged it into `star` under the same name as in `qgraph` (big5):

```{r}
library(star)
data(big5, package = "star")
```

We can combine both data sets to visualize the problems they have in common:

```{r}
d <- bind_rows(Eugene = neo, Dutch = big5, .id = "group")
```

## prepare 

I will use librarian for installing/loading packages. First line checks if librarian is installed, if not, it is installed.

```{r}
if (!require('librarian')) install.packages('librarian')
librarian::shelf(tidyr, dplyr, ggplot2, cowplot, furrr)
```

## values / ideas

```{r}
o6 <- paste0("O", seq(28, 238, 30))
o5 <- paste0("O", seq(23, 233, 30))
sum_score <- d %>% select(all_of(o5)) %>% rowSums(na.rm = TRUE)
sum_score_o6 <- d %>% select(all_of(o6)) %>% rowSums(na.rm = TRUE)

df <- data.frame(o6 = sum_score_o6, o5 = sum_score, group = d$group)
df$Sample <- stringr::str_to_sentence(df$group)
values <- ggplot(df, aes(o6, o5, group = Sample, color = Sample, shape = Sample), alpha = 0.25) + 
  geom_point(alpha=0.5) + 
  geom_smooth(fill = "grey80") + 
  scale_x_continuous("Sum score of facet values") +
  scale_y_continuous("Sum score of facet ideas") +
  scale_color_brewer(type = "qual", palette = "Dark2") + 
  theme_bw() +
  theme_classic() + 
  theme(legend.position = c(0.85, 0.15))
values
cairo_pdf("plots/values.pdf", width = 5, height = 4)
values
dev.off()
```

```{r}
jitter <- 0.05
```

## n96

```{r}
d2 <- d %>%
  group_by(group) %>%
  #filter(group == "eugene") %>%
  select(starts_with("N")) %>%
  mutate(N96 = ifelse(group == "Dutch", N96+jitter, N96-jitter)) %>%
  rowwise() %>%
  mutate(c_sumscore = sum(c_across(N1:N236)))

cor(d2$c_sumscore[d2$group == "Eugene"], d2$N96[d2$group == "Eugene"], use = "complete.obs")
cor(d2$c_sumscore[d2$group == "Dutch"], d2$N96[d2$group == "Dutch"], use = "complete.obs")

n96 <- ggplot(d2, aes(N96, c_sumscore,
               group = group, color = group, shape = group)) + 
  stat_summary(fun.data = "mean_cl_normal",
              fun.args = list(conf.int = .95)) + 
 stat_summary(fun.data = "mean_cl_normal",
              fun.args = list(conf.int = .95), geom = "line", linetype = "dashed") + 
  scale_color_brewer(type = "qual", palette = "Dark2", guide = "none") + 
  scale_shape(guide = "none") + 
  theme_classic() +
  scale_y_continuous("Sum score neuroticism")
n96
```

## C35

Item C35 violated monotonicity in both data sets. As can be seen in Figure x the main problem is at scale point 1. ...

For the dutch data set, the relationship is even negative (r = -0.09). Taking voting seriously is slightly negatively associated with conscientiousness. The item is clearly problematic.


```{r}
d2 <- d %>%
  group_by(group) %>%
  select(starts_with("C")) %>%
  mutate(C35 = ifelse(group == "Dutch", C35+jitter, C35-jitter)) %>%
  rowwise() %>%
  mutate(c_sumscore = sum(c_across(C5:C235), !c_across(C35)))
table(big5[, "C35"])
table(neo_pi_r_eugene[, "C35"])

c35plot <- ggplot(d2, aes(C35, c_sumscore,
               group = group, color = group, shape = group)) + 
  stat_summary(fun.data = "mean_cl_normal",
              fun.args = list(conf.int = .95)) + 
 stat_summary(fun.data = "mean_cl_normal",
              fun.args = list(conf.int = .95), geom = "line", linetype = "dashed") + 
  theme_classic() +
  scale_color_brewer(type = "qual", palette = "Dark2", guide = "none") + 
  scale_shape(guide = "none") + 
  scale_y_continuous("Sum score conscientiousness")

c35plot
d2 %>%
  group_by(group) %>%
  summarize(cor(C35, c_sumscore, use = "complete.obs"))
```

## E17

```{r}
a5 <- paste0("E", seq(27, 237, 30)) # c(seq(22, 234, 30), 
d2 <- d %>%
  filter(!is.na(E17)) %>%
  group_by(group) %>%
  mutate(E17 = ifelse(group == "Dutch", E17+jitter, E17 - jitter)) %>%
  select(a5, E17) %>%
  rowwise() %>%
  mutate(c_sumscore = sum(c_across(a5), na.rm = T))

e17plot <- ggplot(d2, aes(E17, c_sumscore, 
                          group = group, color = group, shape = group)) + 
stat_summary(fun.data = "mean_cl_normal",
              fun.args = list(conf.int = .95)) + 
 stat_summary(fun.data = "mean_cl_normal",
              fun.args = list(conf.int = .95), geom = "line", linetype = "dashed") + 
  scale_y_continuous("Sum score cheerfulness") +
  scale_color_brewer(type = "qual", palette = "Dark2", guide = "none") + 
  scale_shape(guide = "none") + 
  theme_classic()
e17plot
d2 %>%
  group_by(group) %>%
  summarize(cor(E17, c_sumscore, use = "complete.obs"))
```


## A164

```{r}
a5 <- paste0("A", seq(24, 234, 30))
d2 <- d %>%
  # for scale point there is only 1 answer!
  filter(!is.na(A164), A164 >= 2) %>%
  group_by(group) %>%
  select(a5, A164) %>%
  mutate(A164 = ifelse(group == "Dutch", A164+jitter, A164-jitter)) %>%
  rowwise() %>%
  mutate(c_sumscore = sum(c_across(a5), na.rm = T))
table(big5[, "A164"])
table(neo_pi_r_eugene$A164)

d2$Sample <- d2$group
a164plot <- ggplot(d2, aes(A164, c_sumscore,
               group = Sample, color = Sample, shape = Sample)) + 
 stat_summary(fun.data = "mean_cl_normal",
              fun.args = list(conf.int = .95)) + 
 stat_summary(fun.data = "mean_cl_normal",
              fun.args = list(conf.int = .95), geom = "line", linetype = "dashed") + 
theme_classic() +
  scale_color_brewer(type = "qual", palette = "Dark2") + 
  scale_y_continuous("Sum score modesty") +
  scale_x_continuous(limits = c(1, 5)) + 
  theme(legend.position = c(0.85, 0.85))

a164plot
d2 %>%
  group_by(group) %>%
  summarize(cor(A164, c_sumscore, use = "complete.obs"))
```

## A144

```{r}
a5 <- paste0("A", seq(4, 214, 30))
d2 <- d %>%
  filter(!is.na(A144)) %>%
  group_by(group) %>%
  mutate(A144 = ifelse(group == "Dutch", A144+jitter, A144-jitter)) %>%
  select(a5, A144) %>%
  rowwise() %>%
  mutate(c_sumscore = sum(c_across(a5), na.rm = T))

a144plot <- ggplot(d2, aes(A144, c_sumscore,
               group = group, color = group)) + 
 stat_summary(fun.data = "mean_cl_normal",
              fun.args = list(conf.int = .95)) + 
theme_classic() +
  scale_color_brewer(type = "qual", palette = "Dark2") + 
  scale_y_continuous("Sum Score Trust")

a144plot

d2 %>%
  group_by(group) %>%
  summarize(cor(A144, c_sumscore, use = "complete.obs"))
```

## combine

Combine all plots

```{r}
p <- plot_grid(plotlist = list(a164plot, c35plot, e17plot, n96), ncol = 2, labels = "AUTO", label_x = 0.52)
cairo_pdf("plots/part2.pdf", width = 16*0.65, height = 10*0.65)
p
dev.off()
p
```
